Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-02-01 komlos.tex
gamma proof

Rademacher complexity \cite{Bartlett:2003:RGC:944919.944944} is a known method to estimate the generalization ability of an empirical risk minimizer (ERM) of a family of function over a finite set of i.i.d.\ sampled points. We aim here to find an analog for actively selected, rather than i.i.d.\ sampled, points. \cite{langberg2010universal, tolochinsky2018coresets} and references within tie the coreset complexity to the VC dimension of the family function, and the average sensitivity of the dataset. However, these relationships come up more as tools for constructing a coreset rather than a single complexity measure aimed to characterize some generalization ability. We point out the connection between the discrepancy of a function family, informally defined as the result of a two player game, one choosing a set of signs and the other a function from the family, and coreset complexity. Is is known that small discrepancy implies an offline coreset construction algorithm  (see e.g.\ \cite{phillips2009small}, Theorem 1.1). In this paper we point out this connection and provide an algorithm taking use of a small discrepancy bound to construct a coreset in the streaming setting, as well as the offline setting.

@@

These definition are intentionally very similar to those of the Rademacher complexity $R_m(\F)$.
The only difference is that we consider the minimal value over $\sigma$ rather than its expectation for a uniform random $\sigma$.
Rademacher complexity uses random signs because it measures the effect of data being chosen uniformly at random (or drawn from an unknown distribution). See \cite{Bartlett:2003:RGC:944919.944944} for other notions of complexity and relationships between them.\footnote{\cite{Bartlett:2003:RGC:944919.944944} use the term ``discrepancy'' to define a different quantity than the one above.}
We claim that $D_m$ is the right measure of complexity for bounding the size of coresets (coreset complexity) and for creating mergeable sketches (streaming coreset complexity).
Trivially,  $D_m \le R_m$. 
We show, however, that $D_m(\F) = o(R_m(\F))$ for a wide range of interesting problems in machine learning.
This immediately proves the existence of small coresets for these problems.

@@

\begin{corollary}
The class discrepancy of the Gaussian kernel $K(q,x) = \exp(-\gamma \|x-q\|^2)$ in dimension $d$ is $O((1+\gamma)\exp(\gamma)\sqrt{d}/m)$.
\end{corollary} 
This improves upon the recent result of \cite{DBLP:journals/corr/abs-1802-01751} by proving the existence of $\eps$ approximation corsets of size $\sqrt{d}/\eps$ for Gaussian kernel density, in the case where $\gamma$ is constant. 
This also resolves the open problem raised by \cite{DBLP:journals/corr/abs-1802-01751} and matches their lower bound.   
For non-constant $\gamma$ assume w.l.o.g. $\|q-x\| \le 1$. The Taylor series of the Gaussian kernel $K$ exhibits $|\alpha_k| \le \gamma^k/k!$.
Plugging into the equation in the proof of Theorem~\ref{thm:analytic2} we get that the sum determining the constant is upper bounded by
$\sum_{k=1}^\infty \gamma^{k}(k\log^3(k))^{1/2}/k! = O\left((1+\gamma) \exp(\gamma)\right)$.

>>>>

Rademacher complexity \cite{Bartlett:2003:RGC:944919.944944} is a known method to estimate the generalization ability of an empirical risk minimizer (ERM) of a family of function over a finite set of i.i.d.\ sampled points. We aim here to find an analog for actively selected, rather than i.i.d.\ sampled points. \cite{langberg2010universal, tolochinsky2018coresets} and references within tie the coreset complexity to the VC dimension of the family function, and the average sensitivity of the dataset. However, these relationships come up more as tools for constructing a coreset rather than a single complexity measure aimed to characterize some generalization ability. We point out the connection between the discrepancy of a function family, informally defined as the result of a two player game, one choosing a set of signs and the other a function from the family, and coreset complexity. Is is known that small discrepancy implies an offline coreset construction algorithm  (see e.g.\ \cite{phillips2009small}, Theorem 1.1). In this paper we point out this connection and provide an algorithm taking use of a small discrepancy bound to construct a coreset in the streaming setting, as well as the offline setting.

@@

%These definition are intentionally very similar to those of the Rademacher complexity $R_m(\F)$.
%The only difference is that we consider the minimal value over $\sigma$ rather than its expectation for a uniform random $\sigma$.
%Rademacher complexity uses random signs because it measures the effect of data being chosen uniformly at random (or drawn from an unknown distribution). See \cite{Bartlett:2003:RGC:944919.944944} for other notions of complexity and relationships between them.\footnote{\cite{Bartlett:2003:RGC:944919.944944} use the term ``discrepancy'' to define a different quantity than the one above.}
%We claim that $D_m$ is the right measure of complexity for bounding the size of coresets (coreset complexity) and for creating mergeable sketches (streaming coreset complexity).
%Trivially,  $D_m \le R_m$. 
%We show, however, that $D_m(\F) = o(R_m(\F))$ for a wide range of interesting problems in machine learning.
%This immediately proves the existence of small coresets for these problems.

@@

\begin{corollary}
For cases where $\|q-x\| \leq 1$ for all $q,x$, the class discrepancy of the Gaussian kernel $K(q,x) = \exp(-\gamma \|x-q\|^2)$ in dimension $d$ is $O(\gamma\exp(\gamma)\sqrt{d}/m)$.
\end{corollary} 
This improves upon the recent result of \cite{DBLP:journals/corr/abs-1802-01751} by proving the existence of $\eps$ approximation corsets of size $\sqrt{d}/\eps$ for Gaussian kernel density, in the case where $\gamma$ is constant. 
This also resolves the open problem raised by \cite{DBLP:journals/corr/abs-1802-01751} and matches their lower bound.   
For non-constant $\gamma$ assume w.l.o.g. $\|q-x\| \le 1$. The Taylor series of the Gaussian kernel $K$ exhibits $|\alpha_k| \le \gamma^k/k!$.
Plugging into the equation in the proof of Theorem~\ref{thm:analytic2} we get that the sum determining the constant is upper bounded by
$$\sum_{k=1}^\infty \gamma^{k}(k\log^3(k))^{1/2}/k! = O\left(\sum_{k=1}^\infty \gamma^{k}/(k-1)!\right) = O\left(\gamma \exp(\gamma)\right)$$
